---
title: "SQL"
output: html_document
---

## Filtering Rows and Columns 

With SQL, you can filter columns with SELECT and rows with WHERE. Let's look at an example from the Lahman baseball database. First, we need to load some packages. 

```{r warning=FALSE,message=FALSE}
library(Lahman)
library(sqldf)
```

Example 1: Suppose we wanted to see the homerun totals from the New York yankees 
```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting WHERE teamID='NYA' and yearID=1927"
sqldf(query)
```

Example 2: Suppose we waned to see players on the New York Yankees who have atleast 40 homeruns
```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting WHERE teamID='NYA' and HR>=40"
sqldf(query)
```

Example 3: Suppose we wanted to see players who have hit more than 40 homeruns with less than 60 strikeouts 
```{r}
query<-"SELECT playerID,teamID,yearID,HR,SO FROM Batting WHERE HR>40 and SO<60"
sqldf(query)
```

## Order By

With SQL, we can also use Order By to sort information. Below are a few examples in which we used Order By. 


Example 1: Suppose we wanted to see players who have more than 50 homeruns, where the order goes from top to bottom according to highest homerun total 
```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting 
WHERE HR>50 
ORDER BY HR DESC"
sqldf(query)
```

Example 2: Suppose we wanted to see players who have batted atleast 400 times while striking out less than 10 times where the order descends with the most strike outs at the top
```{r}
query<-"SELECT playerID,yearID,teamID,SO,HR,AB FROM Batting 
WHERE SO<10 and AB>=400
ORDER BY SO DESC"
sqldf(query)
```

##Aggregation

With SQL, we can also combine information into one group. Below are a few examples of how we used aggregation. 

Example 1:Suppose we wanted to see the sum all of Babe Ruths homeruns 
```{r}
query<-"SELECT playerID, sum(HR) FROM Batting 
WHERE playerID='ruthba01'
GROUP BY playerID"
sqldf(query)
```

Example 2:Suppose we wanted to see the sum of all players homeruns who have atleast 600 homeruns in descending order 
```{r}
query<-"SELECT playerID, sum(HR) FROM Batting 
GROUP BY playerID
HAVING sum(HR)>=600
ORDER BY HR DESC"
sqldf(query)
```

Example 3: Suppose we wanted to see all players whose average homeruns is greater than 30 in descending order
```{r}
query<-"SELECT playerID, avg(HR) FROM Batting 
GROUP BY playerID
HAVING avg(HR)>30
ORDER BY avg(HR) DESC"
sqldf(query)
```



