---
title: "SQL"
output: html_document
---

## Filtering Rows and Columns 

With SQL, you can filter columns with SELECT and rows with WHERE. Let's look at an example from the Lahman baseball database. First, we need to load some packages. 

```{r warning=FALSE,message=FALSE}
library(Lahman)
library(sqldf)
```

Example 1: Suppose we wanted to see the homerun totals from the New York yankees 
```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting WHERE teamID='NYA' and yearID=1927"
sqldf(query)
```

Example 2: Suppose we waned to see players on the New York Yankees who have atleast 40 homeruns
```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting WHERE teamID='NYA' and HR>=40"
sqldf(query)
```

Example 3: Suppose we wanted to see players who have hit more than 40 homeruns with less than 60 strikeouts 
```{r}
query<-"SELECT playerID,teamID,yearID,HR,SO FROM Batting WHERE HR>40 and SO<60"
sqldf(query)
```

## Order By

With SQL, we can also use Order By to sort information. Below are a few examples in which we used Order By. 


Example 1: Suppose we wanted to see players who have more than 50 homeruns, where the order goes from top to bottom according to highest homerun total 
```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting 
WHERE HR>50 
ORDER BY HR DESC"
sqldf(query)
```

Example 2: Suppose we wanted to see players who have batted atleast 400 times while striking out less than 10 times where the order descends with the most strike outs at the top
```{r}
query<-"SELECT playerID,yearID,teamID,SO,HR,AB FROM Batting 
WHERE SO<10 and AB>=400
ORDER BY SO DESC"
sqldf(query)
```

##Aggregation

With SQL, we can also combine information into one group. Below are a few examples of how we used aggregation. 

Example 1:Suppose we wanted to see the sum all of Babe Ruths homeruns 
```{r}
query<-"SELECT playerID, sum(HR) FROM Batting 
WHERE playerID='ruthba01'
GROUP BY playerID"
sqldf(query)
```

Example 2:Suppose we wanted to see the sum of all players homeruns who have atleast 600 homeruns in descending order 
```{r}
query<-"SELECT playerID, sum(HR) FROM Batting 
GROUP BY playerID
HAVING sum(HR)>=600
ORDER BY HR DESC"
sqldf(query)
```

Example 3: Suppose we wanted to see all players whose average homeruns is greater than 30 in descending order
```{r}
query<-"SELECT playerID, avg(HR) FROM Batting 
GROUP BY playerID
HAVING avg(HR)>30
ORDER BY avg(HR) DESC"
sqldf(query)
```


##Joins

With Joining, we can make tables based on common data in order to make more data output options

Example 1: Suppose we want to see the first and last name, along with team, year, and homeruns for Babe Ruth
```{r}
query<-"SELECT nameFIRST,nameLAST,teamID,yearID,HR 
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE Batting.playerID='ruthba01'"
sqldf(query)
```

Example 2: Suppose we wanted to find all instances of players getting more than 50 homeruns starting with the greatest number. Include first and last names, team, year, and home runs.
```{r}
query<-"SELECT nameFIRST,nameLAST,teamID,yearID,HR
FROM Batting INNER JOIN Master
On Batting.playerID=Master.playerID
WHERE HR>50
ORDER BY HR DESC"
sqldf(query)
```

Example 3: Suppose we wanted to find Babe Ruth's homeruns by season, with the team name displayed.
```{r}
query<-"SELECT playerID,name,Batting.yearID,Batting.HR
FROM Batting INNER JOIN Teams
ON Batting.teamID=Teams.teamID and Batting.yearID=Teams.yearID
WHERE playerID='ruthba01'"
sqldf(query)
```

Example 4: Suppose we wanted to find Babe Ruth's homerun totals, with first and last name and the full team name displayed.

```{r}
query<-"SELECT nameFIRST,nameLAST,name,Batting.yearID,Batting.HR 
FROM (Batting INNER JOIN Master
ON Batting.playerID=Master.playerID) INNER JOIN Teams 
ON Batting.teamID=Teams.teamID and Batting.yearID=Teams.yearID
WHERE Batting.playerID='ruthba01'"
sqldf(query)
```

Example 5: Suppose we wanted to find career homerun totals for players who have at least 600 homeruns. Display from the highest amount to the lowest, with the first and last names.
```{r}
query<-"SELECT nameFIRST,nameLAST,sum(HR)
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
GROUP BY Batting.playerID
HAVING sum(HR)>=600
ORDER BY sum(HR) DESC"
sqldf(query)
```

Example 6: Suppose we wanted to list all players, first and last name, with the first name of Bob who averaged more than one million dollars salary over their career.
```{r}
query<-"SELECT nameFIRST,nameLAST,avg(salary)
FROM Salaries INNER JOIN Master
ON Salaries.playerID=Master.playerID
GROUP BY Salaries.playerID
HAVING nameFIRST='Bob' and avg(salary)>1000000
ORDER BY avg(salary) DESC"
sqldf(query)
```

